---
title: "Aviation Analytics"
author: "Aankur Bhatia"
date: '`r Sys.Date() `'
output: html_notebook
---

First task is to load and tidy up the data, before that we load the libraries

```{r}
#global options set
options(digits = 2, scipen = 5)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(forcats)
library(extrafont)
library(extrafontdb)
library(knitr)
library(DT)
library(DMwR)
library(mlr)

#load the NTSB data set
aviation=read.csv("AviationDataEnd2016UP.csv",na.strings = "",stringsAsFactors = TRUE)
str(aviation)
```

let us now summarize the dataset 

```{r}
summary(aviation)
```

convert event date and publication date to date objects

```{r}
head(aviation)
aviation$Event.Date=ymd(aviation$Event.Date)
aviation$Publication.Date=dmy(aviation$Publication.Date)
```

We would like to focus the analysis on the US as 94% observations are from the US

Remove all other data that is not from the US Country

```{r}
aviation=aviation %>% 
  filter(Country=="United States")
```

Here are the top questions that we would like to ask

1. In the US, has aviation accidents causing fatality improved in the past 70 years?

2. Which phase (ex. take off, cruise, landing..) of the flight contributed to the most fatality? 

3. What was the purpose of the flight that contributed the most fatalities? 

4. Which of the 50 states in the US saw the most fatal accidents? 

```{r eval=TRUE, message=FALSE, echo=FALSE}
# we do the following function
#1. check for duplicates
aviation=distinct(aviation)
#check for column wise missing data
#we can use summarize column function of mlr
summarizeColumns(aviation)
#alternativelya
apply(X = aviation, function(x) sum(is.na(x)),MARGIN = 2)
```

remove rows that have more than 30% NAs

```{r}
manyna=manyNAs(aviation,nORp = 0.3)
#1255 rows - we remove them
aviation=aviation[-manyna,]
```

separate the data into year, month, day

```{r}
aviation$year=year(aviation$Event.Date)
aviation$month=month(aviation$Event.Date,label = TRUE,abbr = TRUE)
aviation$day=day(aviation$Event.Date)
```

Drop columns that have more than 30% NAs in the data 

```{r}
columns=as.data.frame(apply(X = aviation, function(x) sum(is.na(x)),MARGIN = 2))
columns=columns %>% 
  mutate(cname=row.names(columns))
```

